---
title: "EoAR with Zeros"
author: "Trent McDonald"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_notebook: default
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: inline
---

In this vignette, I demonstrate the that EoAR models 
give the same mortality estimate regardless of the 
number of zero sites included. In other words, that EoAR 
naturally accomodates for a large numbers of zeros. 

My example 
data is the bat mortality dataset from the EoAR manuscript.

```{r data, message = FALSE, warning = FALSE}
library(tidyverse)
library(EoAR)
data("batMortData", package="EoAR")
batMortData <- batMortData %>%
  mutate(NLBA = 0) # no carcasses found
set.seed(3382001)
#knitr::kable(batMortData)
```

# G distribution

```{r prepG, fig.cap="Figure 1: Distribution of $g$ used in all examples"}

# Compute one average g 
gParams <- EoAR::mixtureBeta(
  batMortData$gAlpha, 
  batMortData$gBeta, 
  batMortData$Turbines)

# increase width of g distribution to illustrate the point
# If we don't do this, narrowness of g makes changes 
# in uncertainty difficult to see
gParams$alpha <- gParams$alpha / 1000
gParams$beta <- gParams$beta / 1000
gParams

plot(seq(.001,.2,length=500), dbeta(seq(.001,.2,length=500), gParams$alpha, gParams$beta), type = "l", ylab="Density of g", xlab="g")
```


# One Site

The following EoAR model sums zeros over all sites and averages
$g$'s. Effectively, the EoAR model uses one site and one data point. 



```{r oneZero}
# Input dataset
zeroData <- batMortData %>% 
  summarize(NLBA = sum(NLBA),
            Turbines = sum(Turbines))

fit <- EoAR::eoar(NLBA ~ 1,
                       beta.params = gParams, 
                       data = zeroData, 
                       offset = NULL, 
                       nadapt = 10000, 
                       nburns = 50000,
                       niters = 50000, 
                       nthins = 100,
                       nchains = 3,
                       quiet = TRUE)
oneZero <- data.frame(
  nSites = 1,
  offset = "no offset",
  model = paste("Y ~",as.character(fit$call$lambda)[3]), 
  mMedian = fit$estimates[labels(fit,"Mtot"),"Estimate"],
  mMean = mean(unlist(fit$out[,"Mtot"])),
  mHat.lo = fit$intervals[labels(fit,"Mtot"),"5%"],
  mHat.hi = fit$intervals[labels(fit,"Mtot"),"95%"]
)

summary(fit)
oneZero
```
# Two Sites

```{r twoZeros}
# Input dataset
zeroData <- data.frame(
  NLBA = rep(0,2),
  Turbines = rpois(2, 500)
)

fit <- EoAR::eoar(NLBA ~ 1,
                       beta.params = gParams, 
                       data = zeroData, 
                       offset = NULL, 
                       nadapt = 10000, 
                       nburns = 50000,
                       niters = 50000, 
                       nthins = 100,
                       nchains = 3,
                       quiet = TRUE)
twoZero <- data.frame(
  nSites = 2,
  offset = "no offset",
  model = paste("Y ~",as.character(fit$call$lambda)[3]), 
  mMedian = fit$estimates[labels(fit,"Mtot"),"Estimate"],
  mMean = mean(unlist(fit$out[,"Mtot"])),
  mHat.lo = fit$intervals[labels(fit,"Mtot"),"5%"],
  mHat.hi = fit$intervals[labels(fit,"Mtot"),"95%"]
)

summary(fit)
twoZero
```

# Five Sites
```{r fiveZeros}
# Input dataset
zeroData <- data.frame(
  NLBA = rep(0,5),
  Turbines = rpois(5, 500)
)

fit <- EoAR::eoar(NLBA ~ 1,
                       beta.params = gParams, 
                       data = zeroData, 
                       offset = NULL, 
                       nadapt = 10000, 
                       nburns = 50000,
                       niters = 50000, 
                       nthins = 100,
                       nchains = 3,
                       quiet = TRUE)
fiveZero <- data.frame(
  nSites = 5,
  offset = "no offset",
  model = paste("Y ~",as.character(fit$call$lambda)[3]), 
  mMedian = fit$estimates[labels(fit,"Mtot"),"Estimate"],
  mMean = mean(unlist(fit$out[,"Mtot"])),
  mHat.lo = fit$intervals[labels(fit,"Mtot"),"5%"],
  mHat.hi = fit$intervals[labels(fit,"Mtot"),"95%"]
)

summary(fit)
fiveZero
```

# One Hundred Sites

```{r hundredZeros}
# Input dataset
zeroData <- data.frame(
  NLBA = rep(0,100),
  Turbines = rpois(100, 500)
)

fit <- EoAR::eoar(NLBA ~ 1,
                       beta.params = gParams, 
                       data = zeroData, 
                       offset = NULL, 
                       nadapt = 10000, 
                       nburns = 50000,
                       niters = 50000, 
                       nthins = 100,
                       nchains = 3,
                       quiet = TRUE)
hundredZero <- data.frame(
  nSites = 100,
  offset = "no offset",
  model = paste("Y ~",as.character(fit$call$lambda)[3]), 
  mMedian = fit$estimates[labels(fit,"Mtot"),"Estimate"],
  mMean = mean(unlist(fit$out[,"Mtot"])),
  mHat.lo = fit$intervals[labels(fit,"Mtot"),"5%"],
  mHat.hi = fit$intervals[labels(fit,"Mtot"),"95%"]
)

summary(fit)
hundredZero
```

# Two Thousand Sites

```{r twothousandZeros}
# Input dataset
zeroData <- data.frame(
  NLBA = rep(0,2000),
  Turbines = rpois(2000, 500)
)

fit <- EoAR::eoar(NLBA ~ 1,
                       beta.params = gParams, 
                       data = zeroData, 
                       offset = NULL, 
                       nadapt = 10000, 
                       nburns = 50000,
                       niters = 50000, 
                       nthins = 100,
                       nchains = 3,
                       quiet = TRUE)
twothousandZero <- data.frame(
  nSites = 2000,
  offset = "no offset",
  model = paste("Y ~",as.character(fit$call$lambda)[3]), 
  mMedian = fit$estimates[labels(fit,"Mtot"),"Estimate"],
  mMean = mean(unlist(fit$out[,"Mtot"])),
  mHat.lo = fit$intervals[labels(fit,"Mtot"),"5%"],
  mHat.hi = fit$intervals[labels(fit,"Mtot"),"95%"]
)

summary(fit)
twothousandZero
```

# Final Comparison of Results
```{r comparision}
ans <- rbind(oneZero, twoZero, fiveZero, hundredZero, twothousandZero)

knitr::kable(ans)
```
