---
title: "EoAR With Zeros And Offsets"
author: "Trent McDonald"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_notebook: default
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: inline
---

In this vignette, I demonstrate EoAR models fitted 
to no targets, but I include an offset. 

My example 
data is the bat mortality dataset from the EoAR manuscript.

```{r data, message = FALSE, warning = FALSE}
library(tidyverse)
library(EoAR)
data("batMortData", package="EoAR")
batMortData <- batMortData %>%
  mutate(NLBA = 0) # no carcasses found
set.seed(82001)
#knitr::kable(batMortData)
```

# G distribution

```{r prepG, fig.cap="Figure 1: Distribution of $g$ used in all examples"}

# Compute one average g 
gParams <- EoAR::mixtureBeta(
  batMortData$gAlpha, 
  batMortData$gBeta, 
  batMortData$Turbines)

# increase width of g distribution to illustrate the point
# If we don't do this, narrowness of g makes changes 
# in uncertainty difficult to see
gParams$alpha <- gParams$alpha / 1000
gParams$beta <- gParams$beta / 1000
gParams
```

# One Site, One Turbine

```{r oneZero}
# Input dataset
zeroData <- data.frame(
  NLBA = rep(0,1),
  Turbines = 2000
)

fit <- EoAR::eoar(NLBA ~ 1,
                       beta.params = gParams, 
                       data = zeroData, 
                       offset = NULL, 
                       nadapt = 10000, 
                       nburns = 50000,
                       niters = 50000, 
                       nthins = 100,
                       nchains = 3,
                       quiet = TRUE)
r1 <- data.frame(
  nSites = nrow(zeroData),
  offset = 1,
  model = paste("Y ~",as.character(fit$call$lambda)[3]), 
  mMedian = fit$estimates[labels(fit,"Mtot"),"Estimate"],
  mMean = mean(unlist(fit$out[,"Mtot"])),
  lambdaMean = mean(unlist(fit$out[,"lambda"]))
)

summary(fit)

```

# One Site, 2000 Turbines

```{r oneZeroWithOffset}
fit <- EoAR::eoar(NLBA ~ 1,
                       beta.params = gParams, 
                       data = zeroData, 
                       offset = log(Turbines), 
                       nadapt = 10000, 
                       nburns = 50000,
                       niters = 50000, 
                       nthins = 100,
                       nchains = 3,
                       quiet = TRUE)
r2 <- data.frame(
  nSites = nrow(zeroData),
  offset = zeroData$Turbines,
  model = paste("Y ~",as.character(fit$call$lambda)[3]), 
  mMedian = fit$estimates[labels(fit,"Mtot"),"Estimate"],
  mMean = mean(unlist(fit$out[,"Mtot"])),
  lambdaMean = mean(unlist(fit$out[,"lambda"]))
)

summary(fit)
```

# Five Sites Averaging 400 Turbines

```{r fiveZeroWithOffset}
# Input dataset
zeroData <- data.frame(
  NLBA = rep(0,4),
  Turbines = rpois(4, 400)
)
zeroData <- zeroData %>% 
  bind_rows(data.frame(NLBA = 0, Turbines = 2000 - sum(.$Turbines)))
knitr::kable(zeroData)

fit <- EoAR::eoar(NLBA ~ 1,
                       beta.params = gParams, 
                       data = zeroData, 
                       offset = log(Turbines), 
                       nadapt = 10000, 
                       nburns = 50000,
                       niters = 50000, 
                       nthins = 100,
                       nchains = 3,
                       quiet = TRUE)
r3 <- data.frame(
  nSites = nrow(zeroData),
  offset = mean(zeroData$Turbines),
  model = paste("Y ~",as.character(fit$call$lambda)[3]), 
  mMedian = fit$estimates[labels(fit,"Mtot"),"Estimate"],
  mMean = mean(unlist(fit$out[,"Mtot"])),
  lambdaMean = mean(unlist(fit$out[,"lambda[1]"]))
)

summary(fit)
```

# Twenty Sites Averaging 100 Turbines

```{r twentyZeroWithOffset}
# Input dataset
zeroData <- data.frame(
  NLBA = rep(0,19),
  Turbines = rpois(19, 2000/20)
)
zeroData <- zeroData %>% 
  bind_rows(data.frame(NLBA = 0, Turbines = 2000 - sum(.$Turbines)))
knitr::kable(zeroData)

fit <- EoAR::eoar(NLBA ~ 1,
                       beta.params = gParams, 
                       data = zeroData, 
                       offset = log(Turbines), 
                       nadapt = 10000, 
                       nburns = 50000,
                       niters = 50000, 
                       nthins = 100,
                       nchains = 3,
                       quiet = TRUE)
r4 <- data.frame(
  nSites = nrow(zeroData),
  offset = mean(zeroData$Turbines),
  model = paste("Y ~",as.character(fit$call$lambda)[3]), 
  mMedian = fit$estimates[labels(fit,"Mtot"),"Estimate"],
  mMean = mean(unlist(fit$out[,"Mtot"])),
  lambdaMean = mean(unlist(fit$out[,"lambda[1]"]))
)

summary(fit)
```



# Final Table

```{r}
ans <- rbind(r1, r2, r3, r4)
ans <- ans %>% 
  mutate(lambdaScaled = lambdaMean * nSites * offset)
knitr::kable(ans)
```


Note that *lambdaMean* is mortalities per turbine per year, and is constant across rows with the same number of turbines.  Multiplying mortatlities per turbine by number of turbines is approximately the total number killed (i.e., *lambdaScaled* = $Mtot = T\lambda$ where $T$ = number of turbines). 
